\CatchFileBetweenTags{\AlphaInvVal}{calculations/constants.tex}{AlphaInvVal}
\CatchFileBetweenTags{\VonKlitzingVal}{calculations/constants.tex}{VonKlitzingVal}
\CatchFileBetweenTags{\AlphaInvExperimentalValue}{calculations/constants.tex}{AlphaInvExperimentalValue}

\begin{abstract}
We present Informational Energetics (IE), a unified framework for modeling the persistence of complex adaptive systems against entropy. Any system that persists must solve a universal control problem: maintaining structural coherence against environmental fluctuations. By reinterpreting control theory through the lens of information theory and thermodynamics, we derive the Universal Architecture of Persistence: a recursive, six-dimensional set of constraints (Capacity, Map, Protocol, Governor, Toll, and Margin).

To validate the framework's universality, we apply IE to the most fundamental persistent system: the vacuum. Standard physics treats its properties (dimensionality, gauge symmetries, constants) as axiomatic inputs. We reverse this order: treating the vacuum as a persistent system, subject to the same architectural constraints that govern all persistent systems. This enforces a strict sequence of architectural constraints that eliminates all but a single, unique solution. This paper establishes the theoretical foundation for the $E_8$-Persistence Theory by deriving the architectural requirements for the vacuum as a persistent system: Finiteness (preventing divergence), Unitarity (conserving information), and Causality (enforcing temporal order).

We show that these constraints strictly mandate a positive-definite, even, self-dual lattice identifying an $E_8$ lattice projected onto $D=4$ spacetime as the unique geometric solution, characterized by the derived invariant integers $\mathbb{S} =$ $\{\Delta{=}43\text{ (temporal)},$ $\nu{=}16\text{ (chiral)},$ $\sigma{=}5\text{ (interaction)},$ $\chi{=}2\text{ (topological)}\}$.

The $E_8$-Persistence Theory makes a falsifiable prediction: All physical constants derive from this geometry with zero free parameters. We validate this by deriving the Fine-Structure Constant ($\alpha^{-1}$) as the Geometric Impedance of the vacuum. The Von Klitzing Constant ($R_K$) and Elementary Charge ($e$) follow as geometric consequences.
\begin{equation}
\alpha^{-1} = \underbrace{\pi\Delta}_{\text{Capacity}}
+ \,\underbrace{\chi}_{\text{Map}}
- \,\underbrace{\frac{1}{D\Delta - \sigma}}_{\text{Protocol}}
- \,\underbrace{\frac{\chi}{\Delta}}_{\text{Governor}}
+ \,\underbrace{\frac{1}{N^3} \cdot \frac{\chi}{\sigma} \cdot \left( 1 - \frac{\sigma}{D\Delta} \right)}_{\text{Toll}}
+ \,\underbrace{\frac{1}{L_{embed} \cdot (\sigma + 1) \cdot \Delta^2}}_{\text{Margin}}
\end{equation}
Where $L_{embed} = \nu + \sigma + \chi + 2D$ and $N = 2\nu$. Our parameter-free calculation gives $\alpha^{-1} = \AlphaInvVal \dots$, in agreement with CODATA 2022 to within $1.68\sigma$ and with the Morel 2020 value to within $0.58\sigma$.
\end{abstract}