\section{System 0: The Specification, The architectural requirements for persistence: Finiteness, Unitarity, Causality}
\label{sec:System0_vacuum_requirements}
The principles of Informational Energetics, if truly universal, must apply to the most fundamental layer of reality: the vacuum itself. We now perform the act of translation of the universal architecture of persistence to this specific domain to determine the unique configuration that minimizes the Net Entropic Impedance ($Z_{IE}$) to its theoretical floor. In this view, the universe is not a static given, but a self-optimizing physical system subject to the same architectural constraints governing information, energy, and stability necessary for any entity to persist.

The translation of IE's six universal pillars into the specific language of physics yields three non-negotiable requirements for the substrate of reality:
\begin{itemize}
    \item \textbf{Finiteness:} ($\mathrm{CAP}, \mathrm{MAR}$) To prevent energetic and informational divergence.
    \item \textbf{Unitarity:} ($\mathrm{MAP}, \mathrm{PRO}$) To ensure the lossless conservation of information.
    \item \textbf{Causality:} ($\mathrm{GOV}, \mathrm{TOL}$) To enforce a well-defined temporal evolution.
\end{itemize}

The remainder of this section will derive these three properties in detail, building the complete requirements for a persistent universe.

\subsection{The Closure Constraint}
\label{subsec:closure}

Unlike open systems, the vacuum is the totality of existence; it possesses no external environment to provide initial conditions, offload entropy, or store structural memory. IE therefore imposes the \textbf{Closure Constraint}: The substrate of reality must be entirely self-defining. Because the vacuum possesses no external environment, it has no source from which to obtain initial conditions, boundary data, or selection parameters. Its geometry must follow necessarily from its own architectural requirements.

\textit{Any candidate substrate requiring outside data to specify its initial conditions is not a fundamental ground state.}

\subsection{Persistence Requires \texorpdfstring{$\mathrm{CAP}, \mathrm{MAR}$}{CAP, MAR} via Finiteness}
\label{sec:finiteness}
For the vacuum to persist, its definition must be self-consistent. A system defined by infinite properties contains no information and cannot maintain a stable structure. In physics, such inconsistencies manifest as energetic divergences, as exemplified by the Ultraviolet (UV) Catastrophe in standard Quantum Field Theory. IE resolves this by enforcing finiteness as a primary requirement of existence. This corresponds to the \textbf{Capacity} and \textbf{Margin} pillars.

\textbf{Energetic Divergence:} Information is physical and requires energy to store \cite{landauer_irreversibility_1961}. A continuous volume, containing infinite potential information, would therefore possess infinite energy and instantly collapse into a singularity. Under the Closure Constraint (\cref{subsec:closure}), no external mechanism exists to regulate or shed this divergence. To prevent this, the substrate must be \textbf{Discrete}: there must exist a fundamental, indivisible unit of information that sets a maximum density.

\textbf{Algorithmic Complexity:} To minimize Algorithmic Complexity, the substrate must be \textbf{Homogeneous}. In Algorithmic Information Theory, the structural complexity of a random graph scales with its size ($K(S) \propto N$). For the universe to be scalable without additional processing overhead, its Kolmogorov Complexity must be $O(1)$, independent of size (Closure Constraint). This requires that the local topology be the same regardless of location and direction, minimizing the structural description to a single, repeating rule. Consequently, the universe must be a \textbf{Lattice}: a structure that is both discrete (finite) and has a constant structure in every direction (regular), rather than a continuous manifold or a random graph.


\subsection{Persistence Requires \texorpdfstring{$\mathrm{MAP}, \mathrm{PRO}$}{MAP, PRO} via Unitarity}
\label{sec:unitarity}

For the vacuum to persist, it must maintain a stable Map over time. This requires that information is perfectly conserved. By the Closure Constraint (\cref{subsec:closure}), the vacuum must be a perfectly closed and lossless information network, making Unitarity a structural necessity, not merely a mathematical convenience. This maps to the \textbf{Protocol} and \textbf{Map} pillars.

\subsubsection{Lattice Must Be Positive Definite}
\textit{Information-Theoretic Justification}: In a metric space, the norm $|\mathbf{v}|^2 = g_{\mu\nu}v^\mu v^\nu$ represents the information distance from the origin (reference state). For the system to have a well-defined minimum energy configuration (stable ground state), this distance must satisfy:
\begin{equation}
    |\mathbf{v}|^2 \geq 0 \quad \forall \mathbf{v} \neq 0
\end{equation}
A metric with negative eigenvalues (Lorentzian) allows $|\mathbf{v}|^2 < 0$, implying an imaginary ``distance'' from the reference state and preventing the definition of a stable ground state. Furthermore, a Lorentzian metric admits non-trivial null vectors ($|\mathbf{v}|^2 = 0$ where $\mathbf{v} \neq 0$), allowing for the creation of infinite information density without exceeding the capacity budget ($|k\mathbf{v}|^2 = 0$ for any scalar amplitude $k$), which violates the \textbf{Finiteness} pillar. Therefore, the \textbf{substrate} must be Euclidean.

\textbf{Epistemic Note (Substrate vs. Projection):} The Euclidean positive-definiteness derived here applies strictly to the \emph{substrate} lattice. The observed Lorentzian signature $(-,+,+,+)$ emerges dynamically from the causal projection derived later in \cref{sec:metric_signature}. This two-level structure is mathematically consistent: the substrate memory is a static definite metric, while the active processor operates with an indefinite metric distinguishing time from space.

\subsubsection{Lattice Must Be Mappable to a Torus}
\label{sec:torus}
To satisfy \textbf{Finiteness}, the system must be spatially bounded. To satisfy \textbf{Protocol} (Unitarity), it must be closed (no edges). By the combinatorial Gauss-Bonnet theorem, a regular lattice of fixed vertex coordination tiled on a closed surface of Euler characteristic $\chi$ requires topological defects if $\chi \neq 0$. The sphere ($\chi = 2$) therefore requires defects that break translational invariance. The \textbf{Torus} ($T^n$, $\chi=0$) is the unique closed, orientable topology that admits a defect-free regular tiling, simultaneously satisfying Finiteness (bounded), Unitarity (closed), and Homogeneity (flat).

Consequently, the statistical evolution of the vacuum is defined by the Partition Function on a torus\footnote{Here, $\beta$ is the formal periodicity parameter of the imaginary time cycle. We use the partition function formalism for state enumeration on a torus, not to claim the vacuum has a literal temperature.}:
\begin{equation}
    Z(\beta) = \sum_{\text{states}} e^{-S_{\text{config}}} = \text{Tr}(e^{-\beta H})
\end{equation}

\subsubsection{Lattice Must Be Even}
\label{sec:even}
A stable Map requires path independence, which in turn requires evenness. A persistent system demands a unique, unambiguous equilibrium state. If the state count $Z$ depended on the path taken through moduli space (parameterization) rather than the configuration itself, the vacuum would lack a stable \textbf{Map}. 

This requires $Z(\beta)$ to be single-valued under the modular transformation $T: z \to z+1$. The Jacobi Theta Function, $\Theta_\Lambda(z) = \sum e^{i \pi z |\mathbf{v}|^2}$, counts these states. For \textbf{Even} lattices ($|\mathbf{v}|^2 = 2n$), the exponent $2\pi i n z$ is invariant under the shift. However, any odd-norm vector ($|\mathbf{v}|^2 = 2n+1$) introduces a sign inversion ($\Theta \to -\Theta$), rendering the Map multi-valued. Thus, the lattice must be \textbf{Even}.

\subsubsection{Lattice Must Be Self-Dual and Unimodular (Read/Write Symmetry)}
To prevent information loss via destructive interference or Landauer erasure, the encoding operation (write) and the decoding operation (read) must be informationally equivalent. This is enforced by the modular transformation $\mathcal{S}: z \to -1/z$, which maps the lattice to its reciprocal (Fourier dual).

By the Poisson Summation formula, the partition function transforms as:
\begin{equation}
    \Theta_\Lambda(-1/z) \propto \frac{1}{\text{vol}(\Lambda)} \Theta_{\Lambda^*}(z)
\end{equation}
For the Map to remain invariant ($Z_{\Lambda} = Z_{\Lambda^*}$), the lattice must be \textbf{Self-Dual} ($\Lambda = \Lambda^*$). This strictly enforces \textbf{Unimodularity} ($\text{vol}(\Lambda) = 1$), as the only scalar satisfying $V = 1/V$ is $1$. Any other volume introduces an irreversible scaling factor, violating Unitarity.

\textbf{Requirements:} The lattice of the vacuum must be \textbf{Positive Definite, Unimodular, Even, and Self-Dual}.

\subsection{Persistence Requires \texorpdfstring{$\mathrm{GOV}, \mathrm{TOL}$}{GOV, TOL} via Causality}
\label{sec:causality}
For the vacuum to persist, it must not only exist stably but also \textit{evolve} in a well-defined manner. This creates a fundamental information-theoretic challenge: how to project the vast state space of a lattice node onto a single, linear temporal axis without ambiguity. This maps to the \textbf{Governor} and \textbf{Toll} pillars.

\textbf{The Serialization Problem:} From the \textbf{Capacity} pillar, we establish that any persistent system must possess a finite state-space cardinality, denoted $N$ (the number of distinct configurations a single node can occupy). To evolve, the system must serialize these $N$ potential states onto a discrete timeline defined by the \textbf{Temporal Modulus} ($\Delta$), representing the number of available temporal slots in one fundamental causal cycle.

\textbf{Causal Aliasing:} To express its full informational capacity, the system must serialize all $N$ potential states into $\Delta$ discrete temporal slots per fundamental cycle. By the Pigeonhole Principle, if $N > \Delta$, the number of states exceeds the available temporal slots. Consequently, at least one temporal coordinate must simultaneously host multiple distinct states. This results in \textbf{Causal Aliasing}: distinct information states become concurrently superimposed in time, destroying the system's ability to maintain a well-defined serial history.
    
\textbf{The Persistence Inequality:} To enforce a strict arrow of time and satisfy the Causality requirement, the projection must satisfy the Persistence Inequality: $\Delta \ge N$. The temporal container must meet or exceed the information content it holds. This inequality implicitly requires full capacity utilization: every channel in the $N$-dimensional state space must be expressible within the $\Delta$ temporal slots of a fundamental cycle. If any channel were permanently inaccessible, it would represent ``dead'' structure, information capacity that exists but cannot be exercised, which explicitly violates the Finiteness and Closure constraints that the substrate contain no inert auxiliary overhead.

\subsection{Summary: The Architectural Requirement of the Vacuum}
Through the rigorous application of Informational Energetics, we have fully specified the architectural requirements for the substrate of reality:

\begin{enumerate}
    \item ($\mathrm{CAP}, \mathrm{MAR}$) \textbf{Finite Lattice}, required by Finiteness.
    \item ($\mathrm{MAP}, \mathrm{PRO}$) \textbf{Positive Definite, Unimodular, Even, and Self-Dual}, required by Unitarity. 
    \item ($\mathrm{GOV}, \mathrm{TOL}$) Any causal system must satisfy a \textbf{Causal Projection}: a serialization of its $N$-dimensional state-space onto a discrete timeline of length $\Delta \ge N$.
\end{enumerate}

With this specification complete, the physical substrate is no longer an open exploration but a constrained mathematical problem: find the unique geometric object that satisfies all of these requirements simultaneously.