\subsection{Resolution of the Distler-Garibaldi No-Go Theorem}
\label{sec:resolution_of_the_distler_garibaldi}

A critical objection to $E_8$ unification is the theorem by Distler and Garibaldi stating that the group's real fundamental representation ($\mathbf{248}$) precludes chiral fermions in a continuous algebraic embedding. Our framework circumvents this via \textbf{Metric Projection} rather than algebraic subgrouping.

The theorem applies to maps between manifolds of identical signature. However, the physical derivation involves a signature change:
\begin{equation}
    \mathbb{R}^8 \text{ (Euclidean)} \xrightarrow{\mathcal{P}} M^{1,3} \text{ (Lorentzian)}
\end{equation}

\begin{enumerate}
    \item \textbf{Clifford Incompatibility:} The Clifford algebra of the lattice is $Cl(8,0)$, which admits only real (Majorana-Weyl) spinors. The Clifford algebra of spacetime is $Cl(1,3) \cong M_4(\mathbb{C})$, which admits complex Weyl spinors.
    \item \textbf{The Projection Cost:} To map the real lattice data onto the complex manifold without generating non-physical ``ghost states" (negative norm), the projection operator must satisfy the causal constraint. This acts as a \textbf{Chiral Diode}:
    \begin{equation}
        P_{causal} = \frac{1 - \gamma^5}{2}
    \end{equation}
    In standard QFT, $\gamma^5$ is an operator. In this geometric framework, $\gamma^5$ is the volume element of the 4D projection. The ``Mirror" fermions required by $E_8$ symmetry still exist in the lattice mathematical structure, but they are \textbf{geometrically orthogonal} to the observer's timeline. They are not erased; they are causally disconnected.
\end{enumerate}

Thus, the chiral asymmetry of the Standard Model is identified as the artifact of projecting a real, symmetric, Euclidean geometry onto a complex, causal, Lorentzian timeline.