\section{The Geometric Invariants}
\label{sec:SystemI}

To determine the physical architecture of reality, we do not start with observation; we start with the requirements of persistence. We posit that the vacuum self-organizes to maximize its information processing capability. In this section, we identify the unique geometric structure that satisfies the simultaneous thermodynamic and information-theoretic constraints of Informational Energetics.

\subsection{The Standard Model Ansatz}
In standard physics, the dimensionality of spacetime ($D=4$), the existence of chiral fermions, and the specific gauge groups are treated as axiomatic backgrounds, a pre-existing stage upon which fields act. Standard theory offers no structural reason why the universe is 4-dimensional rather than 10-dimensional, why time exists as a distinct metric signature, or why the fine-structure constant has a specific value necessary for structural complexity. These fundamental properties are inputs, not outputs.

\subsection{The System Specification}
We define the Lattice Substrate by instantiating the six pillars of persistence. These five integers constitute the complete input set $\mathbb{S}$ from which the physical universe emerges.

\begin{enumerate}
    \item \textbf{Capacity ($\Delta E$): Fundamental Resonance ($\Delta=43$).} 
    The maximum non-repeating frequency of the lattice (Heegner Number), representing the bit-depth of the vacuum.
    
    \item \textbf{Identity ($\Delta I$): Interaction Symmetry ($\sigma=5$).} 
    The geometric rank required to encode the unified force ($SU(5)$ precursor).
    
    \item \textbf{Protocol ($MI$): Chiral Capacity ($\nu=16$).} 
    The active degrees of freedom available for matter storage (The Weyl Spinor).
    
    \item \textbf{Governor ($G$): Topological Boundary ($\chi=2$).} 
    The Euler characteristic required for closed loops (stable particles), enforcing the finiteness of the field.
    
    \item \textbf{Temporal Cost ($T$): Causality ($-1$).} 
    The metric signature requirement for state updates, creating the arrow of time.
    
    \item \textbf{Persistence Margin ($PM$): Manifold Rank ($D=4$).} 
    The minimum dimensional embedding required to support the projection.
\end{enumerate}

\subsection{Derivation A: The Hardware Selection (Why \texorpdfstring{$E_8$}{E8}?)}
We first determine the geometry of the lattice itself before projecting it into spacetime.


\subsubsection{The Conservation of Information (Self-Duality)}
For a lattice field theory to preserve unitarity over long timescales, it must be indistinguishable from its dual ($\Lambda = \Lambda^*$). 
\begin{itemize}
    \item \textbf{Impedance Match:} Physically, self-duality ensures the lattice possesses unit impedance regarding information flow; there is no geometric scattering or loss when mapping the lattice to its reciprocal.
    \item \textbf{CPT Symmetry:} If $\Lambda \neq \Lambda^*$, the mapping introduces scale-dependent phase artifacts, violating global Charge--Parity--Time symmetry. 
\end{itemize}
Therefore, the substrate must be an \textbf{Even, Self-Dual Lattice}. (``Even'' ensures that all vector norms are even integers, enforcing discrete energy quantization—the lattice supports only bosonic-type excitations).

\subsubsection{Kneser's Theorem}
Mathematical constraint restricts our search space.\cite{kneser_klassenzahlen_1957} Even, self-dual lattices exist uniquely only in dimensions divisible by 8:
\[ D \in \{8, 16, 24, \dots\} \]
This strictly eliminates any lattice solution in dimensions $D<8$.

\subsubsection{The Principle of Minimum Configurational Entropy}
To minimize the Entropic Action of the substrate, the system must not only minimize geometric complexity (dimension) but also eliminate arbitrary selection parameters. We analyze the population of even, self-dual lattices permitted by Kneser's Theorem ($D = 8k$):

\begin{itemize}
    \item \textbf{D=8:} A unique solution exists (The $E_8$ lattice).
    \item \textbf{D=16:} Two distinct solutions exist ($E_8 \oplus E_8$ and $D_{16}^+$).
    \item \textbf{D=24:} Twenty-four distinct solutions exist (The Niemeier lattices).
\end{itemize}

A vacuum established at $D=16$ would possess an irreducible \textbf{Configurational Entropy} of $S_{config} = k_B \ln(2)$, representing the information required to distinguish between the two topological isomers. A vacuum at $D=24$ would have $S_{config} = k_B \ln(24)$.

The $E_8$ lattice ($D=8$) is the unique solution where $S_{config} = k_B \ln(1) = 0$. It is selected not merely for its low dimensionality, but because it is the only self-dual geometry that forms a deterministic ground state without inherent topological ambiguity.

\subsection{Derivation B: The Projection (Why \texorpdfstring{$D=4$}{D=4} and \texorpdfstring{$\nu=16$}{nu=16}?)}
A static 8-dimensional block cannot process information; processing requires a flow (Input vs. Output). The system must break the $E_8$ symmetry to distinguish the ``Observer" (Spacetime) from the ``System" (Internal States).

\subsubsection{The Symmetric Decomposition (Space vs. Charge)}
The projection must preserve the self-duality property in the subsystems to maintain local conservation. The unique symmetric splitting of $E_8$ is:
\begin{equation}
    E_8 \to D_4 \oplus D_4
\end{equation}

While other decompositions exist (e.g., $E_8 \to A_8$), the $D_4 \oplus D_4$ split is the unique decomposition among maximal rank subgroups that preserves the self-duality of the subspaces \cite{conway_sphere_1988}. By Conway \& Sloane [8] (Ch 4, \S 4.7), other maximal subgroups (e.g., $A_8$, $D_8$) fail to preserve the integer norm condition required for a unitary lattice projection without introducing scaling factors that break self-duality.

This decomposition partitions the 8 dimensions into two orthogonal sectors with distinct physical roles:
\begin{enumerate}
    \item \textbf{Sector A (External Spacetime):} The first $D_4$ lattice defines the coordinate addresses of the lattice nodes. Since $\text{Rank}(D_4)=4$, the observable universe is strictly fixed at \textbf{$D=4$}.
    \item \textbf{Sector B (Internal Symmetry):} The second $D_4$ lattice encodes the internal state (charge, spin, isospin) at each coordinate. These dimensions do not manifest as spatial directions but as the \textbf{Gauge Symmetries} of the Standard Model (detailed in \cref{app:RootInventoryAndStatePartition}).
\end{enumerate}
This structural partition explains why the universe appears 4-dimensional while possessing complex internal forces, without requiring the hidden spatial dimensions of Kaluza-Klein theory.

\subsubsection{The Chiral Capacity (\texorpdfstring{$\nu=16$}{nu=16})}
We must determine the ``Bit-Depth" of the lattice nodes. What is the minimum amount of information required to define a distinct, charged particle?

The $E_8 \to D_4 \oplus D_4$ decomposition creates a local symmetry of $Spin(8)$. However, this structure faces a critical physical limitation:
\begin{enumerate}
    \item \textbf{The Real Constraint:} Due to the unique symmetry structure of Spin(8) (called Triality), the spinor representations in $Spin(8)$ are Real (Self-Conjugate). Mathematically, this means a particle is indistinguishable from its antiparticle ($\psi = \bar{\psi}$). A universe built on this logic would prevent the encoding of \textbf{$U(1)$ charges} (Electromagnetism) or the emergence of Parity Violation.
    
    \item \textbf{The Complex Solution:} To distinguish Matter from Antimatter, the system requires \textbf{Complex Representations} (where $\psi \neq \bar{\psi}$). We must extend the symmetry to the minimal group that supports complex ``Weyl Spinors."
    
    \item \textbf{The Minimal Extension:} The smallest group containing $Spin(8)$ that supports complex representations is \textbf{Spin(10)}. The size of the fundamental data packet (spinor) in this group is:
    \begin{equation}
        \Delta_{\text{spin}} = 2^{5-1} = \mathbf{16}
    \end{equation}
\end{enumerate}
Thus, $\nu=16$ is not an arbitrary particle count; it is the Geometric Bit-Depth required to encode complex, chiral information on the lattice.

\paragraph{The Total State Space ($N$)}
The projection creates two orthogonal sectors (Matter and Mirror) to preserve self-duality. Consequently, the total information capacity of a single lattice node is the sum of these two chiral halves:
\begin{equation}
    N = \nu_{\text{matter}} + \nu_{\text{mirror}} = 16 + 16 = \mathbf{32}
\end{equation}
This integer $N=32$ represents the total number of distinct state channels that must be mapped onto the temporal dimension without collision. To prevent aliasing (where distinct states map to the same time coordinate), the fundamental resonance must satisfy $\Delta > N$, establishing the lower bound for the lattice frequency.

\subsection{Derivation C: The Operating System (Metric \& Time)}
Having established a 4D manifold with 16-channel capacity, we must derive the metric signature. The metric signature determines the algebraic structure of the spinor representation.

\subsubsection{The Spinor Constraint}
A 4-dimensional manifold can admit a Euclidean $(+,+,+,+)$ or Lorentzian $(-,+,+,+)$ signature. The choice is determined by which algebra supports the required $\nu=16$ complex states.

\begin{itemize}
    \item \textbf{Euclidean ($4,0$):} The Clifford algebra is $Cl(4,0) \cong \mathbb{H}(2)$ ($2\times2$ Quaternionic matrices). Quaternionic spinors are mathematically ``Real" (Symplectic). They lack the commutative imaginary unit $i$ required to encode quantum phases ($e^{i\theta}$) or distinguish chiral states.
    
    \item \textbf{Lorentzian ($3,1$):} The Clifford algebra is $Cl(3,1) \cong \mathbb{C}(4)$ ($4\times4$ Complex matrices). The introduction of the negative metric signature naturally generates the complex structure. This supports \textbf{Weyl Spinors} with 8 complex components (16 real degrees of freedom), exactly matching the hardware capacity.
\end{itemize}

\textbf{Conclusion:} The requirement for complex information processing ($\nu=16$) mandates a Lorentzian metric. Physically, \textbf{Time} ($ds^2 < 0$) is the geometric cost required to generate the imaginary unit ($i$) in the algebra.

\subsubsection{The Chiral Diode (The Arrow of Time)}

To enforce causal order, the lattice must distinguish ``Past'' from ``Future.'' The $E_8$ root system possesses a total capacity of $N=32$ degrees of freedom, decomposing symmetrically as:
\begin{equation}
    E_8 \to D_4 \oplus D_4 \quad \Rightarrow \quad N = \nu_L + \nu_R = 16 + 16 = 32
\end{equation}
However, a fully symmetric channel allows information to flow bidirectionally (standing waves), which prohibits the formation of a temporal gradient.

To enforce the Persistence Principle, the system undergoes \textbf{Chiral Truncation}. The active state space is restricted to the left-handed sector ($\nu_L = 16$), while the right-handed sector ($\nu_R = 16$) is projected out. This asymmetry acts as a \textbf{Geometric Diode}, only signals propagating ``with the grain'' of the chiral projection are permitted to carry state updates.

\textbf{Physical Consequences (The Pillar Mapping):}
\begin{enumerate}
    \item \textbf{Arrow of Time (Temporal Cost $T = -1$):} The diode enforces unidirectional information flow. This maps to the negative metric signature dimension, acting as the scalar update index. Movement here is irreversible (entropic burn).
    
    \item \textbf{Spatial Capacity (Persistence Margin $PM = +3$):} The remaining orthogonal dimensions provide the volumetric capacity for topological storage. This maps to the positive metric signature, creating the ``Space'' where knots (Identity) reside.
    
    \item \textbf{Parity Violation:} The weak interaction (the mechanism of state change) couples exclusively to left-handed particles. The ``return path'' through right-handed states is geometrically blocked, preventing acausal loops.
    
    \item \textbf{Channel Saturation:} With exactly half the lattice capacity locked out, there is no available bandwidth for supersymmetric partners or additional generations beyond $n_{gen} = \sigma - \chi = 3$.
\end{enumerate}

This structural truncation is the geometric origin of CP violation (Paper III) and the sterility of right-handed neutrinos (Paper II).

\subsection{Derivation D: The System Logic (\texorpdfstring{$\sigma$}{sigma} and \texorpdfstring{$\chi$}{chi})}
We derive the rank of the interaction symmetry ($\sigma$). This derivation is supported by two converging lines of evidence: one from Group Theory (Algebraic) and one from Manifold Geometry (Geometric).

\subsubsection{The Interaction Symmetry (\texorpdfstring{$\sigma=5$}{sigma=5})}
\begin{enumerate}
    \item \textbf{Algebraic Necessity (The Container):} The gauge group must be the minimal simple Lie group capable of embedding the Standard Model groups $SU(3) \times SU(2) \times U(1)$. By representation theory, this requires a group of Rank $\ge 4$. The minimal simple group satisfying this is $SU(5)$, which operates on a fundamental representation of dimension 5.
    
    \item \textbf{Geometric Necessity (The Degrees of Freedom):} Physically, a persistent state is defined by its \textbf{Location} and its \textbf{Boundary}. The minimal embedding vector space $V$ must span:
    \begin{itemize}
        \item \textbf{Spatial Freedom ($D-1=3$):} The 3 dimensions required to define translation (Where is the particle?).
        \item \textbf{Topological Freedom ($\chi=2$):} The 2 dimensions required to define a closed boundary surface (What defines the ``inside" vs ``outside"?).
    \end{itemize}
    These sectors are orthogonal (translation commutes with deformation), so their dimensionalities add linearly:
    \begin{equation}
        \sigma = \dim(V) = 3 \text{ (Space)} + 2 \text{ (Boundary)} = \mathbf{5}
    \end{equation}
\end{enumerate}
This dual convergence identifies $\sigma=5$ as the inevitable Interaction Order.

\subsubsection{The Topological Boundary (\texorpdfstring{$\chi=2$}{chi=2})}
For a particle to be distinct from the vacuum, its boundary must strictly separate the universe into two disjoint sets: ``Inside'' (The System) and ``Outside'' (The Environment).  We assert that the topological boundary of a persistent particle must be a sphere ($\chi=2$) as the \emph{unique} solution to the \textbf{Binary Partition Constraint}.

We derive this by analyzing the Euler Characteristic for closed, orientable surfaces:
\begin{equation}
    \chi = 2 - 2g
\end{equation}
where $g$ is the genus (number of holes).

\begin{itemize}
    \item \textbf{The Exclusion of $g \geq 1$ (The Leaky Partition):} Any topology with one or more holes (Torus $g=1$, Double Torus $g=2$, etc.) fails to define a strict binary separation.
    \begin{itemize}
        \item \textit{Information-Theoretic Failure:} A genus $g \geq 1$ surface is not simply connected. It supports non-contractible loops—paths that thread through the holes without intersecting the surface. This creates informational ambiguity: field lines can interact with the topology without being ``enclosed,'' rendering the definition of total charge ambiguous (Gauss's Law fails).
        
        \item \textit{Thermodynamic Failure:} By the Gauss-Bonnet theorem ($\int_M K \, dA = 2\pi\chi$), any surface with $g \geq 1$ has $\chi \leq 0$, implying neutral or negative total curvature. Such a surface cannot support net positive internal pressure (energy density) against the vacuum, it would structurally collapse.
    \end{itemize}
    
    \item \textbf{The Uniqueness of $g=0$ (The Sphere):} The sphere is the unique closed surface with $g=0$, yielding $\chi=2$, the maximum possible value. It is the only simply connected topology, ensuring that all loops contract to a point. This forces every field line to be explicitly either contained or excluded, enabling the perfect binary partition of state required for a persistent, distinguishable particle.
\end{itemize}

\paragraph{Physical Consequence: Charge Quantization}
The invariant $\chi=2$ is the \emph{necessary and sufficient} condition for \textbf{charge quantization}. Because $\chi$ must be an integer and $\chi=2$ is the unique maximum for closed surfaces, the charge associated with this topology is discrete. You can have 1 sphere or 2 spheres, but not 1.5 spheres. This geometric constraint allows the continuous lattice field to support discrete, countable units of charge.


\subsection{Derivation E: The Fundamental Resonance (\texorpdfstring{$\Delta=43$}{Delta=43})}
Finally, we derive the fundamental frequency of the lattice. This is the only dynamic integer in the set. It must satisfy three simultaneous filters to support a persistent universe.

\subsubsection{Filter 1: Unitarity (The ``History" Constraint)}
For a quantum system to preserve information (Unitarity), the evolution of states must be reversible. In a lattice based on integer arithmetic, this requires that every composite state must decompose into prime components in exactly one way (Unique Factorization).
\begin{itemize}
    \item \textbf{The Problem:} If the mathematical field defining the lattice has a ``Class Number" $h > 1$, unique factorization fails (e.g., $6 = 2 \times 3$ but also $6 = a \times b$). The system loses the history of how the state was formed.
    \item \textbf{The Constraint:} The resonant number $\Delta$ must belong to the special set of integers where $h=1$. These are the Heegner Numbers.
    \item \textbf{Search Space:} $\{1, 2, 3, 7, 11, 19, 43, 67, 163\}$. (By the Stark-Heegner Theorem, no others exist).
\end{itemize}

\subsubsection{Filter 2: Causality (The ``Bandwidth" Constraint)}
The system must map the total capacity of the lattice ($N = 32$ channels) onto the timeline defined by $\Delta$.

\begin{itemize}
    \item \textbf{The Problem:} If the timeline cycle ($\Delta$) is shorter than the number of channels ($N$), the Pigeonhole Principle forces two distinct states to map to the same time coordinate. This creates ``Causal Aliasing'' (Signal Collision), violating the Generalized Nyquist Limit derived in  \cref{app:DerivationOfTheCausalityConstraint}.
    
    \item \textbf{The Constraint:}: To ensure every state has a unique address, we require $\Delta > N = 32$. In the statistical mechanics of the lattice, this acts as the \textbf{Causality Potential} ($V_C$), assigning infinite energy cost to any configuration where temporal aliasing occurs (see  \cref{app:partition-function}).
    
    \item \textbf{Remaining Candidates:} $\{43, 67, 163\}$.
\end{itemize}

\subsubsection{Filter 3: Structural Solvency (The ``Stability" Constraint)}
The resonant frequency determines the stiffness of the vacuum ($\alpha^{-1} \approx \pi\Delta$). This sets the binding energy of all matter. We test the remaining candidates against the Landauer Limit (the minimum energy $k_B T \ln 2$ required to protect information from thermal noise).

\begin{itemize}
    \item \textbf{Candidate $\Delta=163$ ($\alpha \approx 1/514$):} \textbf{Too Weak.} Binding energies drop to $\sim 0.3$ eV. At room temperature, thermal noise would rip electrons off atoms. Matter dissolves into plasma. \textbf{(Eliminated).}
    
    \item \textbf{Candidate $\Delta=67$ ($\alpha \approx 1/212$):} \textbf{Marginal.} Binding energies are $\sim 1.2$ eV. While simple atoms persist, the Signal-to-Noise Ratio is too low to support the complex error-correction required for long-term persistence\footnote{For $\Delta=67$: SNR $\approx 60$. For $\Delta=43$: SNR $\approx 200$. Reliable error correction generally requires SNR $\gtrsim 100$.}. The system is structurally brittle. \textbf{(Eliminated).}
    
    \item \textbf{Candidate $\Delta=43$ ($\alpha \approx 1/137$):} \textbf{Solvent.} Binding energies are $\sim 4$ eV. This creates a deep enough energy well to protect matter from noise ($SNR \gg 1$) while remaining shallow enough to allow dynamic state transitions. It is the unique solution.
\end{itemize}
\textbf{Result:} $\Delta = 43$ is the unique integer solution.



\subsection{Derivation F: The Geometric Origin of Spin}
\label{sec:spin_derivation}

Standard physics categorizes particles by intrinsic angular momentum (Spin), but offers no structural reason why matter is spin-$\frac{1}{2}$ and force carriers are spin-1. In the $E_8$-Persistence framework, spin is identified as the \textbf{Geometric Rank} of the coupling to the lattice substrate.

\subsubsection{Spin-\texorpdfstring{$\frac{1}{2}$}{12}: Topological Nodes (Fermions)}
Fermions occupy the chiral lattice nodes ($\nu=16$). The node topology is closed ($\chi=2$), imposing binary occupancy (0 or 1), a single node cannot store two identical quantum states. This topological constraint manifests as the \textbf{Pauli Exclusion Principle}.

The spinor phase property $\psi(2\pi) = -\psi(0)$ arises from the projection geometry: to sample the full 32-channel capacity of the lattice ($N$) from the 16-channel chiral projection ($\nu$), a rotation must traverse the manifold twice ($720^\circ$). This double-cover structure identifies fermions as ``half-integer'' excitations of the geometry.

\subsubsection{Spin-1: Network Links (Gauge Bosons)}
Gauge bosons act as connections between nodes, coupling to the vector indices of spacetime ($D=4$). As transmission signals rather than storage addresses, they do not occupy the topological boundary ($\chi_{\text{boson}}=0$ relative to nodes), permitting unbounded occupancy—the geometric origin of \textbf{Bose-Einstein statistics}.

\subsubsection{Spin-2: Bulk Geometry (Gravitons)}
Gravity represents deformation of the lattice substrate itself (System VI). It couples to the metric tensor $g_{\mu\nu}$ (rank-2), corresponding to the stress-energy distribution across the bulk volume. The graviton emerges as the Goldstone mode of broken translational symmetry in the presence of matter ($T^{\mu\nu} \neq 0$).

\subsubsection{The Prohibition of Higher Spins}
The $D=4$ lattice geometry strictly prohibits fundamental particles with spin $> 2$:
\begin{itemize}
    \item \textbf{Spin-$\frac{3}{2}$ (Gravitino):} Would require rank-3/2 coupling (vector-spinor), which cannot be constructed from the $D_4 \oplus D_4$ decomposition without introducing mirror fermions forbidden by the chiral truncation (Section IV.F.2).
    \item \textbf{Spin-3 and higher:} Would couple to rank-$n \geq 3$ tensors, which exceed the dimensional capacity of the 4D manifold.
\end{itemize}
This geometric constraint falsifies supersymmetry and higher-spin extensions of the Standard Model.



\subsection{Conclusion: The Invariant Set}
We have successfully derived the complete set of inputs $\mathbb{S} = \{D=4, \Delta=43, \nu=16, \sigma=5, \chi=2\}$ without reference to experimental tuning. 
Mathematically, we identify this set not merely as a geometric preference, but as the \textbf{Gibbs State} of the $E_8$ lattice, the unique vacuum configuration that minimizes the Entropic Action defined by the Hamiltonian $H_{info}$ (detailed in \ref{app:partition-function}). The Standard Model is the thermodynamic ground state of this invariant set. These are the unique \textbf{Characteristic Integers} of a self-dual information processing substrate satisfying persistence, unitarity, and causality constraints within the framework of even lattices and simple Lie groups.
    
\subsection{The Derived Capacities (System Output)} \label{sec:SystemicCapacities}
Before proceeding to the coupling calculations (System II), we must define the total bandwidth available to the system based on the derived invariants. We distinguish between the \textit{informational content} of the lattice and the \textit{persistence budget} required to sustain a state, and the \textit{static load} of the background geometry.

\begin{enumerate}
    \item \textbf{The Systemic Channel ($H_{sys}$):} The sum of the active degrees of freedom available for internal information storage (Chiral + Interaction + Boundary).
    \begin{equation}
        \label{eq:hsys}
        H_{sys} = \nu + \sigma + \chi = 16 + 5 + 2 = \mathbf{23}
    \end{equation} 
    
    \item \textbf{The Full Persistence Budget ($H_{full}$):} The total operating cost for a persistent structure. This adds the Spacetime Embedding Cost to the internal channel. \textit{The Spinor Cost:} Unlike a vector, a spinor must be rotated by $720^\circ$ (4$\pi$) to return to its original state (the spinor double cover). Consequently, the system must pay the existence cost for each of the $D$ dimensions \textbf{twice} to fully define an embedded state.
    \begin{equation}
        \label{eq:hfull}
        H_{full} = H_{sys} + 2D = 23 + 8 = \mathbf{31}
    \end{equation}

    \item \textbf{The Structural Overhead ($H_{struct}$):} The total geometric load required to support the fundamental resonance within the manifold. This combines the spacetime projection cost of the resonance ($\Delta \cdot D$) with the chiral storage capacity ($\nu$). This value represents the static background load against which vacuum potentials (like the Higgs VEV) must be normalized.
    \begin{equation}
        \label{eq:Hstruct}
        H_{struct} = (\Delta \cdot D) + \nu = (43 \times 4) + 16 = \mathbf{188}
    \end{equation}
\end{enumerate}

\subsubsection{The Bandwidth Limit (\texorpdfstring{$c_{eff}$}{ceff})}
Because the lattice is discrete, information propagates via neighbor-to-neighbor updates. The maximum rate of causality is bounded by the lattice node spacing ($\ell$) divided by the update frequency ($\Delta$).

Consequently, Lorentz Invariance is identified not as a geometric axiom, but as the \textbf{Shannon Channel Capacity ($C_{max}$)} of the substrate. The speed of light $c \equiv 1$ represents the maximum bandwidth of the vacuum protocol.
\begin{equation}
    c_{eff} = \frac{\text{Lattice Spacing}}{\text{Update Cycle}} \equiv 1
\end{equation}
This enforces the causality constraint derived in Appendix B: any signal attempting to propagate faster than the protocol allows results in temporal aliasing (Causal Violation) and is filtered by the Causality Potential ($V_C$).
