\section{System 1: The Projection, The \texorpdfstring{$E_8 \rightarrow D_4 \oplus D_4$}{E8toD4D4} geometric projection that uniquely satisfies the System 0 requirements}
\label{sec:system1_projection}

Having established the requirements of the substrate we now identify the unique mathematical structure that satisfies these constraints.

\subsection{The Lattice Selection (\texorpdfstring{$E_8$}{E8})}
\label{sec:lattice_selection}

We will begin by determining the possible dimensions of the lattice, then identify a unique lattice, and finally determine its projection into spacetime.

\subsubsection{The \texorpdfstring{$n \equiv 0 \pmod 8$}{neq0mod8} Constraint}
\label{sec:kneserstheorem}
The requirement of an even, self-dual lattice is remarkably restrictive. A fundamental constraint from the theory of modular forms and lattices\cite{conway_sphere_1988}, states that even, self-dual lattices only exist in dimensions that are multiples of 8. This immediately eliminates the majority of dimensionalities leaving only: 

\[ n \in \{8, 16, 24, \dots\} \]
This strictly eliminates any fundamental lattice solution in dimensions $n<8$. This yields a profound physical consequence: the observed $n=4$ spacetime cannot mathematically host a unitary, self-dual information substrate. The universe we observe cannot be the fundamental hardware; it must be an emergent projection of a higher-dimensional structure.

Furthermore, this excludes $n=10$ (Superstring Theory) as a standalone geometric substrate. Because $10 \not\equiv 0 \pmod{8}$, a 10-dimensional lattice cannot satisfy \textit{Unitarity} (the requirement for an even, self-dual lattice). To recover these properties, String Theory must introduce auxiliary structures (e.g., Calabi-Yau compactifications or appended internal spaces). The selection of which compactification manifold to use is not determined by the theory's intrinsic constraints, requiring external configuration memory and directly violating the Closure Constraint (\cref{subsec:closure}).

\subsubsection{The Principle of Algorithmic Determinism}
\label{sec:principleofalgorithmicdeterminism}

To satisfy \textbf{Map} and \textbf{Unitarity}, the substrate geometry must be intrinsic and self-defining. It cannot depend on arbitrary selection parameters or external boundary conditions. We demonstrate that the constraints on the possible dimensions eliminate all but one unique option:

\begin{enumerate}
    \item \textbf{Unitarity} mandates even, self-dual, unimodular lattices. By Kneser's theorem, such lattices exist only for $n \equiv 0 \pmod{8}$.
    
    \item By the \textbf{Closure Constraint} (\cref{subsec:closure}), the vacuum admits no external configuration memory. Formally, the Information/Closure constraint acts as an exact Kolmogorov complexity filter. The Specification Complexity for the substrate must satisfy $K(\text{Geometry} \mid \text{IE constraints}) = 0$ additional bits. 
    
    \item At $n=8$: The $E_8$ lattice is the \textbf{unique} even self-dual lattice \cite{conway_sphere_1988}. Selecting this geometry from the set of valid $n=8$ solutions requires $\log_2(1) = \mathbf{0 \text{ bits}}$ of external selection information.
    
    \item At $n=16$: Two non-isomorphic solutions exist ($E_8 \oplus E_8$ and $D_{16}^+$). Selecting between them requires $\log_2(2) = \mathbf{1 \text{ bit}}$ of information. Because the vacuum admits $0$ bits of external configuration memory (Closure Constraint), this $1$-bit selection cannot be realized: no physical mechanism exists within the substrate to source this information. 
    
    \item At $n=24$: Twenty-four distinct Niemeier lattices exist, requiring $\log_2(24) \approx \mathbf{4.58 \text{ bits}}$ of selection information, similarly violating the Closure Constraint.
    
    \item The number of distinct even self-dual lattices grows extremely rapidly with dimension. For example, in dimension 32, there are more than 80 million such lattices.
\end{enumerate}

The $E_8$ lattice is selected not merely for its low dimensionality, but because it is the \textbf{only self-dual geometry that requires zero bits of selection information}. It is uniquely self-defining. This differs from the $\Delta$ derivation, where surjection and continuum limit are forced by causal continuity and atomicity (physical requirements) not arbitrary choices between equivalents.

This unique self-consistent solution is selected prior to the emergence of time, eliminating any argument of spontaneous symmetry breaking which would require fluctuations, transition rates, and a dynamical framework that does not yet exist and which the substrate must first generate.

\subsection{The Projection (\texorpdfstring{$D=4$}{D=4} and \texorpdfstring{$\nu=16$}{nu=16})}
\label{sec:projection}

The $E_8$ lattice cannot process information by itself. Processing requires a flow (Input vs. Output). The projection must break the $E_8$ symmetry to distinguish the ``Observer'' (Spacetime) from the ``System'' (Internal States).

\subsubsection{The Symmetric Decomposition (Space vs. Charge)}
\label{sec:symmetric_decomposition}

The projection must be compatible with the self-duality of the parent $E_8$ lattice to maintain local probability conservation. Because $E_8$ is self-dual ($\Lambda = \Lambda^*$), any symmetric decomposition into orthogonal sublattices ($E_8 \to L_4 \oplus L_4$) must admit a specific glue group whose coset structure exactly reconstructs the self-dual parent. To avoid unbalancing the Unitarity requirement, the state space must be partitioned into two symmetric sectors that can support this exact quotient structure.

\textbf{Candidate Rank-4 Subsectors:}
We evaluate rank-4 root sublattices that can symmetrically embed in $E_8$. The irreducible rank-4 root systems are $A_4$, $B_4$, $C_4$, $D_4$, and $F_4$. We restrict our search to sublattices admitting a symmetric embedding $E_8 \to L_4 \oplus L_4$. $B_4$ and $C_4$ are not simply-laced; their root lattices contain roots of two distinct lengths, violating the Evenness requirement. $F_4$, while even, is not self-dual. This leaves $A_4 \oplus A_4$ and $D_4 \oplus D_4$ as the only mathematically viable candidates.

\begin{itemize}
    \item \textbf{The $A_4$ No-Go (Encoding Mismatch):} The $A_4 \oplus A_4$ subsystem has an index of 5 in the $E_8$ lattice. In an information-theoretic framework, its quotient group ($\mathbb{Z}_5$) requires mapping a 5-state system onto binary hardware channels. Local hardware encoding of a 5-element group requires a discrete register of $\lceil \log_2(5) \rceil = 3$ bits, but this necessarily introduces $2^3 - 5 = 3$ phantom codewords with no geometric referent. Under the Closure Constraint, the substrate admits no extraneous structural data; these unphysical states would constitute auxiliary configuration memory, explicitly violating the requirement that the vacuum be entirely self-defining.
    
    \item \textbf{The $D_4$ Solution (The Unitary Partition):} The $D_4 \oplus D_4$ orthogonal decomposition has an index of 4 in $E_8$. Because 4 is a perfect power of two ($\log_2(4) = 2$ exact bits), its quotient group ($\mathbb{Z}_2 \times \mathbb{Z}_2$) can be encoded perfectly without fractional loss, satisfying Unitarity. Furthermore, $D_4$ (the root lattice of $\mathfrak{so}(8)$) is the unique rank-4 even lattice that supports Triality. This provides the exact geometric symmetry required to subsequently construct the Map and Causality pillars (\cref{sec:chiralcapacity}).
\end{itemize}

Therefore, $E_8 \to D_4 \oplus D_4$ is the \textbf{strictly unique symmetric decomposition} that preserves global Unitarity while naturally generating the chiral structure required for information processing.

This unique decomposition partitions the 8 dimensions into two orthogonal sectors with distinct roles:

\begin{enumerate}
    \item \textbf{Sector A (External Spacetime):} The first $D_4$ lattice defines the coordinate addresses of the lattice nodes. Since the $D_4$ root system has rank 4 (four linearly independent simple roots), it naturally projects onto a 4-dimensional coordinate manifold, fixing \textbf{D=4} for spacetime.  The rank equals the number of independent Cartan generators, which specify the coordinate addresses; hence rank 4 implies $D=4$.
    \item \textbf{Sector B (Internal Symmetry):} The second $D_4$ lattice encodes the internal configuration of each node, its state within the substrate's symmetry structure. These four dimensions are not spatial directions but internal degrees of freedom. 
\end{enumerate}

This structural partition explains why the substrate must eventually appear 4-dimensional while possessing complex internal forces, without requiring the hidden spatial dimensions of Kaluza-Klein theory.

\subsubsection{The Bit-Depth of the Node (\texorpdfstring{$\nu = 16$}{nu16})}
\label{sec:chiralcapacity}

To determine the information capacity (Bit-Depth) of the lattice nodes we ask: what is the minimal geometric structure required to define a persistent, distinguishable signal on the lattice?

\textit{Geometric Translation:} In a lattice, information states are encoded as \textbf{orientations}, directional configurations that can be rotated and transformed. 

The mathematical structure governing these rotations in an $n$-dimensional space is the \textbf{Spin group} $Spin(n)$, which describes how states change under geometric transformations while preserving their essential properties.

The $D_4$ lattice spans 4 spatial dimensions, but its associated Lie algebra ($\mathfrak{so}(8)$) acts on 8-dimensional fundamental and spinor representations. As established in \cref{sec:projection}, each $D_4$ factor in the $E_8 \to D_4 \oplus D_4$ decomposition therefore gives a local continuous rotation symmetry of $Spin(8)$. However, this structure faces critical information-theoretic limitations that prevent it from satisfying the \textbf{Map} and \textbf{Causality} pillars.

\begin{enumerate}
    \item \textbf{The Map Constraint (Complex Phase):} 
    While $Spin(8)$ possesses chiral spinors (Left and Right), they are mathematically \textbf{Real} (Majorana-Weyl). In signal processing terms, this means a state vector is identical to its conjugate ($\psi = \bar{\psi}$). A system built on this logic cannot distinguish a signal from its inverse (Phase Ambiguity) because it lacks complex phase information. To support a stable Map, the system requires \textbf{Complex Representations} ($\psi \neq \bar{\psi}$), allowing for the encoding of phase distinct from amplitude.
    
    \item \textbf{The Causality Constraint (Chirality):} 
    To support the \textbf{Causality} pillar (Arrow of Time), the system must distinguish ``Input'' from ``Output.'' Geometrically, this requires \textbf{Chirality}, the ability to distinguish Left-handed projections from Right-handed projections. Odd-dimensional rotation groups (such as $Spin(9)$) admit complex representations but are inherently achiral, failing this requirement.

    \item \textbf{The Minimal Geometric Group:} 
    We seek the minimal geometric group rank that satisfies both existential conditions:
    \begin{itemize}
        \item \textbf{Complex} (to encode Phase/Map)
        \item \textbf{Chiral} (to encode Flow/Causality)
    \end{itemize}
\end{enumerate}

Because the $Spin(8)$ chiral spinors lack complex structure, and $Spin(9)$ complex spinors lack chirality, the unique minimal extension that simultaneously supports Complex, Chiral representations (Weyl-Dirac spinors) is $Spin(10)$. 

Crucially, this does not imply that the $D_4$ substrate geometrically embeds a $Spin(10)$ gauge group. The substrate system strictly prohibits this, as $Spin(10)$ (rank 5) exceeds the structural capacity of the internal $D_4$ sector (rank 4). Rather, $Spin(10)$ strictly defines the \textbf{informational payload size} required to encode a persistent signal. The size of this fundamental data packet is given by the Weyl spinor dimension formula $\nu = 2^{n/2-1}$ for $n$-dimensional chiral spinors:

\begin{equation}
    \nu = 2^{\frac{10}{2}-1} = 2^4 = \mathbf{16}
\end{equation}

To transmit this 16-channel payload without exceeding the rank-4 internal budget, the node utilizes the representations natively available to its $D_4$ ($\mathfrak{so}(8)$) geometry. Because $D_4$ possesses Triality, it supports independent 8-dimensional representations (e.g., $8_L$ and $8_R$). A single node's \textbf{internal symmetry sector} provides exactly $8 + 8 = 16$ channels of hardware capacity, perfectly matching the payload requirement.

Thus, $\nu = 16$ is established as the fundamental bit-depth of the vacuum's hardware; it is the minimal capacity necessary to resolve phase ambiguity and enforce causality, an architectural requirement satisfied by the $Spin(10)$ spinor dimensions.

\subsection{The Total Node Capacity (\texorpdfstring{$N$}{N})}
\label{sec:total_node_capacity}
We derive the total information capacity of a single lattice node, which constrains the state space of all subsequent 
derivations.

The projection $E_8 \to D_4 \oplus D_4$ creates two orthogonal, symmetric subsystems. As derived above, the internal sector requires a bit-depth of $\nu=16$. Because the projection must be compatible with the \textbf{Self-Duality} of the parent $E_8$ substrate, the Spacetime sector must remain informationally symmetric to the Internal sector. Because information capacity (bit-depth) is additive across independent orthogonal sub-spaces, the total channel capacity $N$ is simply the sum of the two sectors:
\begin{equation}
    N = \nu_{\text{Sector A}} + \nu_{\text{Sector B}} = 16 + 16 = \mathbf{32}
\end{equation}

This integer $N = 32$ represents the total number of distinct state channels available for encoding information on the lattice substrate.

\subsection{Dimensional Reduction: Geometric Origin of The Metric Signature and Causality}
\label{sec:metric_signature}

\subsubsection{The Requirements}
To map the 16-channel signal onto a 4-dimensional manifold without loss or ambiguity, the metric algebra must satisfy two strict encoding constraints:
\begin{enumerate}
    \item \textbf{Complex Compression (Map):} The 16 real channels must be compressed into 8 complex degrees of freedom ($16\mathbb{R} \to 8\mathbb{C}$). This requires the algebra to naturally generate a geometric imaginary unit ($i$) to encode phase information.
    \item \textbf{Chiral Sorting (Causality):} The system must distinguish ``Input'' from ``Output'' to enforce the arrow of time. This requires the algebra to support orthogonal projectors ($P_L, P_R$) that split the signal into directed halves ($8\mathbb{C} \to 4\mathbb{C}_L + 4\mathbb{C}_R$).
\end{enumerate}

\subsubsection{The Search Space}
A 4-dimensional manifold admits two fundamental metric signatures. We test each against the requirements:

\begin{itemize}
    \item \textbf{Option A: Euclidean Metric $(+,+,+,+)$.} 
    In a space where all dimensions are spatial, the volume element (pseudoscalar $\omega$) squares to positive unity ($\omega^2 = +1$). 
    \item \textbf{Option B: Lorentzian Metric $(-,+,+,+)$.} 
    In a space with one temporal dimension, the Clifford algebra volume element (the pseudoscalar $\omega$) squares to negative unity ($\omega^2 = -1$).
\end{itemize}

\subsubsection{The Unique Solution}
The Euclidean metric fails the encoding test. Because $\omega^2 = +1$, the center of the Clifford algebra is strictly real ($\mathbb{R} \oplus \mathbb{R}$). It cannot generate the global geometric imaginary unit $i$ required for complex phase compression, nor can it support complex chiral projectors. A universe with this signature would be a static, real-valued block with no capacity for phase or flow.

The \textbf{Lorentzian Metric} is the unique valid solution: $(1,3)$ is equivalent to $(3,1)$ by reordering, and $(2,2)$ fails because $\omega^2 = +1$ in split signature, restoring the real algebra. Because $\omega^2 = -1$, the Clifford volume element functions identically as the geometric imaginary unit ($i \equiv \omega$). This naturally generates the complex structure required to compress the signal ($\nu=16$) and the chiral structure required to sort it.

Crucially, the metric projection assigns the Lorentzian signature $(-,+,+,+)$ exclusively to Sector A (Spacetime). Sector B (Internal Symmetry) retains the original Euclidean signature $(+,+,+,+)$ of the $E_8$ algebra. This asymmetry is not imposed but is the direct consequence of the Causality requirement: only one $D_4$ factor can support a timelike generator, and that assignment uniquely distinguishes the spacetime sector from the internal sector.

\textbf{Conclusion:} The signature $(-,+,+,+)$ is not an arbitrary choice; it is the only algebraic structure capable of processing the system's information content. Physically, the negative sign of Time ($dt^2 < 0$) is the geometric cost required to generate the imaginary unit $i$.

\paragraph{Physical Consequence:}
Once these algebraic structures are established to satisfy the encoding requirement, they manifest in physics as fundamental properties of matter. The \textbf{Matter/Antimatter} distinction emerges from the complex conjugation ($\psi \leftrightarrow \bar{\psi}$) enabled by the imaginary unit. \textbf{Chirality} emerges from the orthogonal projectors. These are not postulates, but unavoidable consequences of encoding a 16-channels onto a 4-dimensional Lorentzian manifold


\subsection{The System Logic (\texorpdfstring{$\sigma$=5}{sigma=5} and \texorpdfstring{$\chi$=2}{chi=2})}
\label{sec:the_system_logic}

The \textbf{Finiteness} and \textbf{Unitarity} pillars require the vacuum to be homogeneous (no preferred positions) and self-dual (informationally symmetric). For a node to possess both a \textbf{coordinate address} and a \textbf{distinguishable Map} while satisfying these constraints, the two specifications must be \textbf{orthogonal}: independent, yet simultaneously well-defined. Geometrically, this requires the substrate to support a \emph{topological defect}, a localized boundary structure capable of separating a region from the vacuum, without specifying what occupies that region. The minimal embedding dimension for this orthogonal decomposition sets the rank of the interaction symmetry ($\sigma$), the geometric degrees of freedom available for force mediation.

\subsubsection{The Topological Boundary (\texorpdfstring{$\chi=2$}{chi=2})}
\label{sec:topological_boundary}
For a topological defect to be distinct from the vacuum, its boundary must strictly separate the universe into two disjoint sets: ``Inside'' (The System) and ``Outside'' (The Environment).  We assert that the topological boundary of a persistent topological defect must be a sphere ($\chi=2$) as the \emph{unique} solution to the \textbf{Binary Partition Constraint}.

We derive this by analyzing the Euler Characteristic for closed, orientable surfaces:
\begin{equation}
    \chi = 2 - 2g
\end{equation}
where $g$ is the genus (number of holes).

\begin{itemize}
    \item \textbf{The Exclusion of $g \geq 1$ (The Leaky Partition):} Any topology with one or more holes (Torus $g=1$, Double Torus $g=2$, etc.) fails to define a strict binary separation.
    \begin{itemize}
        \item \textit{Information-Theoretic Failure:} A genus $g \geq 1$ surface is not simply connected. It supports non-contractible loops, closed paths that thread through the holes without intersecting the surface. This creates informational ambiguity: information flux can traverse the topology without being strictly partitioned into ``Inside'' or ``Outside,'' rendering the total conserved quantity associated with the boundary ill-defined.
        
        \item \textit{Thermodynamic Failure:} By the Gauss-Bonnet theorem ($\int_M K \, dA = 2\pi\chi$), any surface with $g \geq 1$ has $\chi \leq 0$, implying neutral or negative total curvature. Such a surface cannot support net positive entropic pressure (information density) against the bulk substrate: it would structurally collapse.
    \end{itemize}

    \item \textbf{The Uniqueness of $g=0$ (The Sphere):} The sphere is the unique closed surface with $g=0$, yielding $\chi=2$, the maximum possible value. It is the only simply connected topology, ensuring that all loops contract to a point. This forces all information flux to be explicitly either contained or excluded, enabling the perfect binary partition of state required for a persistent, distinguishable topological defect.
\end{itemize}

\paragraph{Forward Implication (System 2 Consequence):}
The invariant $\chi=2$ is the \emph{necessary and sufficient} condition for \textbf{charge quantization}. Because $\chi$ must be an integer and $\chi=2$ is the unique maximum for closed surfaces, the charge associated with this topology is discrete. You can have 1 sphere or 2 spheres, but not 1.5 spheres. This geometric constraint allows the continuous lattice field to support discrete, countable units of charge.

\subsubsection{The Embedding Invariant (\texorpdfstring{$\sigma=5$}{sigma=5})}
\label{sec:embedding_invariant}

We derive $\sigma$ as the minimal coordinate set required to fully specify a persistent entity on the lattice. A localized 
state requires two independent pieces of information: its \textbf{spatial address} (where it is) and its 
\textbf{topological boundary} (what it is). These sectors are orthogonal, one describes position in the manifold, 
the other describes the internal boundary structure, and together they define the minimal embedding dimension:

\begin{itemize}
    \item \textbf{Spatial Freedom ($D_{space} = 3$):} 
    The \textbf{Causality} pillar requires that one dimension of the $D=4$ manifold be allocated to the update stream (Time). This leaves $D-1=3$ dimensions for spatial configuration.

    This value $D_{\text{space}} = 3$ is strictly required by the \textbf{Topological Boundary Constraint}. As derived in \cref{sec:topological_boundary}, the persistent defect must possess an $S^2$ boundary ($\chi=2$). A 2-dimensional spherical boundary cannot be geometrically embedded in a spatial manifold of fewer than 3 dimensions. Therefore, 3D space is the minimal spatial dimension capable of supporting the exact topological defect required for persistence. (In 3D, the second homotopy group $\pi_2(S^2) = \mathbb{Z}$ classifies these stable localized topological defects).
    
    \item \textbf{Topological Boundary ($D_{boundary} = 2$):} The \textbf{Map} pillar requires the topological defect to be distinguished from the vacuum by a closed boundary. As derived in \cref{sec:topological_boundary}, the unique simply-connected solution is the sphere ($S^2$). Regardless of its location in the $3D$ spatial bulk, specifying the internal state of this boundary requires $2$ intrinsic coordinate parameters (e.g., $\theta, \phi$). Because this definition is independent of the defect's spatial position, these $2$ degrees of freedom are informationally orthogonal to $D_{space}$.
\end{itemize}

\textbf{The Orthogonality Condition:}
A fundamental requirement of the \textbf{Homogeneity} defined in System 0 is that the internal definition of a topological defect (its boundary topology) must be independent of its location. 

Mathematically, this requires the total state space to be a \textbf{Direct Product} of the spatial manifold and the internal boundary manifold ($V_{total} = V_{space} \times V_{boundary}$). Because the boundary definition does not depend on position, their vector spaces are orthogonal.

The dimension of the minimal interaction embedding is the sum of these independent sectors:
\begin{equation}
    \sigma = \dim(V_{space}) + \dim(V_{boundary}) = 3 + 2 = \mathbf{5}
\end{equation}

\subsection{The Fundamental Resonance (\texorpdfstring{$\Delta=43$}{Delta=43})}
\label{sec:fundamental_resonance}

We derive the fundamental period of the lattice, the temporal bit-depth of the vacuum. Three independent constraints must be simultaneously satisfied.

\subsubsection{Unitarity and The Unique Factorization Domain Constraint}
\label{sec:fundamental_resonance_filter1}

To apply this Unitarity constraint, we must map the physical period $\Delta$ to the algebraic structure of the causal loop. As proven formally in \cref{app:algebraic_geometry_causal_loop}, a $(1+1)D$ causal sequence on a self-dual lattice forms an elliptic curve. For the evolution operator to be uniquely invertible (Unitary), this curve must possess Complex Multiplication (CM) with a class number of $h=1$. The self-duality of the substrate strictly mandates that the geometric period of the loop ($\Delta$) must equal the algebraic CM discriminant ($D$). 

Consequently, the Unitarity requirement restricts the temporal period $\Delta$ strictly to the Stark-Heegner numbers (the unique solutions for $h=1$): $\{1, 2, 3, 7, 11, 19, 43, 67, 163\}$.

These nine \textbf{Heegner numbers} are the only valid candidates for the temporal resonance of a unitary causal loop.

\subsubsection{Causality (The ``Bandwidth'') Constraint}
\label{sec:fundamental_resonance_filter2}
In \cref{sec:causality} we established that $\Delta > N = 32$ is required to satisfy Causality. Applying this constraint to the set of Heegner numbers leaves:

\begin{equation}
    \Delta \in \{43, 67, 163\}
\end{equation}

\subsubsection{The Self-Clocking Constraint (Causal Continuity)}
\label{sec:fundamental_resonance_filter3}

This constraint imposes a strict upper bound on the cycle length by addressing how the vacuum maintains temporal synchronization.

\begin{itemize}
    \item \textbf{The Mechanism (Signal vs. Overhead):} 
    The fundamental cycle $\Delta$ is not a single indivisible unit, but a periodic causal loop composed of $\Delta$ discrete, atomic temporal steps. These steps partition into $N=32$ active signal states (the informational payload) and $M = \Delta - N$ overhead intervals (structural processing latency where the metric advances but state is static).
    
    \item \textbf{The Self-Clocking Requirement:} 
    By the \textbf{Closure Constraint} (\cref{subsec:closure}), the vacuum possesses no external reference clock. The progression of time must therefore be maintained intrinsically by the lattice's own state transitions. In information theory, a data stream without an external reference is a \textbf{Self-Clocking Signal}.
    
    \item \textbf{The Phase-Lock Limit:} 
    To maintain causal phase-lock without an external clock, empty overhead intervals ($M$) must be informationally anchored by active state transitions ($N$). If the number of empty steps exceeds the number of signal states ($M > N$), by the Pigeonhole Principle, the sequence must contain un-anchored consecutive empty steps. Because these empty steps contain zero distinct informational content, the local node cannot differentiate or count them, resulting in causal drift and the loss of temporal sequence (Acausal Gaps). Furthermore, the node cannot rely on its spatial neighbors to act as an external clock. Because the lattice is translationally homogeneous, all neighboring nodes are geometrically identical during a static interval; they provide no differential information to break the temporal ambiguity.
    \item \textbf{The Bound (Local vs. Global Phase-Lock):} 
    To ensure causal continuity, Informational Energetics mandates that the number of active signal states ($N$) must equal or exceed the number of empty metric updates ($M$). The Finiteness requirement (\cref{sec:finiteness}) mandates that the substrate be \textbf{homogeneous}: its local topology must be identical at every position. Formally, this requires the signal/overhead sequence to be \emph{maximally uniform} (a Balanced Sequence in combinatorics). For a binary sequence of length $\Delta$ with $N$ active states and $M = \Delta - N$ overhead intervals, the Beatty/Fraenkel theorem guarantees the maximum run length of empty steps is $\lceil M / N \rceil$. To prevent unanchored gaps (run length $\le 1$), we require $\lceil (\Delta - N) / N \rceil \le 1$, which strictly enforces the global density bound $N \ge M$. This ensures phase-lock is maintained across all configurations, mathematically equivalent to an RLL$(0,1)$ constraint \cite{immink_runlength-limited_1990}.
    \begin{equation}
        N \ge M \implies N \ge (\Delta - N) \implies \Delta \le 2N
    \end{equation}
\end{itemize}

\textbf{Result:} With $N=32$, the upper limit for a self-clocking fundamental cycle is $\Delta \le 64$. This rigorously eliminates the remaining Stark-Heegner candidates $67$ and $163$, leaving $\Delta = 43$ as the strictly unique geometric solution for the temporal period.

\subsection{Manifold Resolution}
\label{sec:manifold_resolution}

Given that the substrate is a $D=4$ dimensional manifold with fundamental temporal cycle $\Delta=43$, the theoretical maximum number of distinct spacetime slots the local geometry can support per fundamental update cycle is the product:

\begin{equation}
    R_M = D\Delta = 4 \times 43 = 172
\end{equation}

Persistent topological defects occupy subsets of this address space; their operational costs in System 2 scale with the fraction of $R_M$ utilized.

This completes the substrate's information geometry: $\nu=16$ encodes what a single spinor holds, $N=32$ the total channels at a node, and $R_M=172$ the spacetime addresses available per cycle.

\CatchFileBetweenTags{\InvLIntrinsic}{constants.tex}{InvLIntrinsic}
\CatchFileBetweenTags{\InvLEmbed}{constants.tex}{InvLEmbed}
\CatchFileBetweenTags{\InvLSubstrate}{constants.tex}{InvLSubstrate}

\subsection{The Entropic Loads: Resolution Limits of the Vacuum}
\label{sec:lattice_loads}

The vacuum's persistence depends on three hierarchical scales of informational complexity. Each scale defines a resolution limit, a threshold below which signals dissolve into the substrate.

\subsubsection{The Intrinsic Load (\texorpdfstring{$L_{intrinsic} = 23$}{Lintrinsic}): Internal Specification}
The information required to specify a topological defect's internal state comprises three orthogonal components. The chiral configuration ($\nu=16$) encodes phase and handedness. The interaction embedding ($\sigma=5$) encodes the coordinate degrees of freedom: spatial address ($D_{space}=3$) and boundary coordinates ($D_{boundary}=2$, the continuous parameters required to specify a point on the boundary surface). The topological invariant ($\chi=2$) encodes the global identity of the boundary, the discrete classification that distinguishes the sphere ($\chi=2$) from all other closed surfaces, independent of local coordinates. Together these three orthogonal components set the internal resolution limit: the precision with which the vacuum can distinguish \emph{what} a defect is (internal state) from \emph{where} it is (spatial position).
\begin{equation}
    L_{intrinsic} = \nu + \sigma + \chi = 16 + 5 + 2 = \mathbf{23}
\end{equation}

\subsubsection{The Embedding Load (\texorpdfstring{$L_{embed} = \InvLEmbed$}{Lembed}): Manifold Address}
The total information to specify a defect's complete configuration: internal state plus spatial address.
\label{sec:lembed}

Because the chiral signal states are dynamic entities within the $D=4$ manifold, specifying their complete state address requires more than just spatial coordinates. In discrete control theory and state-space representations, uniquely specifying a dynamic state's full trajectory requires establishing both its position vector ($D$ dimensions) and its transition rate or orientation vector ($D$ dimensions). This necessitates a total embedding cost of $2D = 8$ independent degrees of freedom to uniquely specify the dynamic state address on the manifold. (We note that this purely information-theoretic requirement emerges in higher-level physics as Hamiltonian Phase Space and conjugate momentum, though we rely only on the geometric state-space bound here).

\begin{equation}
    L_{embed} = L_{intrinsic} + 2D = 23 + 8 = \mathbf{\InvLEmbed}
\end{equation}

\subsubsection{The Substrate Load (\texorpdfstring{$L_{substrate} = \InvLSubstrate$}{Loverhead})}
The vacuum's intrinsic structural complexity.   This sets the substrate capacity—the total information the vacuum must process per fundamental cycle. 

\begin{itemize}
    \item \textbf{$\nu = 16$:} Channel infrastructure. The spinor information at each site.

    \item \textbf{$R_M (D \cdot \Delta) = 172$:} Manifold Resolution.
\end{itemize}
\begin{equation}
    L_{substrate} = (R_M) + \nu = (4 \times 43) + 16 = \mathbf{\InvLSubstrate}
\end{equation}


\subsection{The \texorpdfstring{$E_8$}{E8}-Persistence System Specification}
\label{sec:system1spec}
We have now identified a unique solution that satisfies the architectural requirements of System 0.

\begin{itemize}
    \item[] \textbf{Substrate: The $E_8$ Manifold.} The $D_4 \oplus D_4$ decomposition of $E_8$ fixes $D=4$ as the unique dimensionality supporting Finiteness and Causality.

    \item \textbf{Capacity ($CAP$): The Operational Envelope ($N=32$).} The $N=32$ chiral channels within each $\Delta=43$ cycle define the total state-transition budget. The clock period sets the envelope; the channel count determines what fills it.

    \item \textbf{Map ($MAP$): The Persistent Embedding ($\sigma=5$).} The sum of spatial ($D_{\text{space}}=3$) and topological ($D_{\text{boundary}}=2$) degrees of freedom required to uniquely embed a persistent entity in the manifold.
    
    \item \textbf{Protocol ($PRO$): The Chiral Bit-Depth ($\nu=16$).} The minimal spinor dimension admitting complex structure and chirality, enabling lossless information propagation. Required by Unitarity for read/write symmetry.

    \item \textbf{Governor ($GOV$): The Topological Boundary ($\chi=2$).} The sphere ($S^2$), unique simply-connected closed surface, enforcing charge quantization and preventing field divergence. Required by Finiteness.

    \item \textbf{Toll ($TOL$): The causal cost of persistence ($\tau = -1$, $M=11$).} Structurally, the Lorentzian metric ($\tau = -1$) encodes time by sacrificing one spatial dimension's definiteness. Operationally, the $M = \Delta - N = 11$ overhead intervals in the fundamental cycle provide the temporal slack required for lossless state transitions.

    \item \textbf{Margin ($MAR$): The Resolution Floor ($L_{embed}=31$).} The total embedding dimension sets the geometric noise floor below which signals dissolve into the substrate. In Quiescent Equilibrium, this floor is \textbf{saturated}: the minimal resolvable signal (the electron mass, derived in Section~\ref{sec:system2_geometric_impedance}) is the proof of solvency. The vacuum persists because its resolution limit is exactly minimally sufficient.
\end{itemize}

\subsection{Conclusion: \texorpdfstring{$E_8$}{E8}-Persistence theory}
The constraints of \textbf{Finiteness}, \textbf{Unitarity}, and \textbf{Causality} do not permit a family of solutions; they converge on a \textbf{unique} geometric structure: the $E_8$ lattice projected onto a causal $D=4$ manifold. This unique geometric structure is accompanied by the set of Characteristic Integers: $\mathbb{S} = \{\Delta{=}43, \nu{=}16, \sigma{=}5, \chi{=}2\}$.

We have established the foundation of the $E_8$-Persistence theory: a unique substrate on which other systems can persist.