\subsection{System Summary: The Hardware Specification}
\label{sec:hardware_specification}

For the purpose of subsequent derivations, we distinguish four derived capacities that define the bandwidth and impedance of this vacuum substrate.

\subsubsection{The Intrinsic Capacity (\texorpdfstring{$H_{sys} = 23$}{Hsys})}
The sum of the active degrees of freedom available for internal information storage (Chiral + Interaction + Boundary):
\begin{equation}
    H_{sys} = \nu + \sigma + \chi = 16 + 5 + 2 = \mathbf{23}
\end{equation}

\subsubsection{The Embedding Budget (\texorpdfstring{$H_{full} = 31$}{Hfull})}
The total cost to sustain a persistent structure within the spacetime manifold. Because spinors live on the double cover of the manifold (requiring $720^\circ$ rotation), the system pays the existence cost for the $D=4$ dimensions twice.
\begin{equation}
    H_{full} = H_{sys} + 2D = 23 + 8 = \mathbf{31}
\end{equation}

\subsubsection{The Structural Impedance (\texorpdfstring{$H_{struct} = 188$}{Hstruct})}
The static load of the vacuum background. This value acts as the normalization factor for Vacuum Expectation Values (VEV), representing the geometric pressure against which potentials must be calibrated.
\begin{equation}
    H_{struct} = (\Delta \cdot D) + \nu = (43 \times 4) + 16 = \mathbf{188}
\end{equation}

\subsubsection{The Lattice Capacity (\texorpdfstring{$N = 32$}{N})}
The total bit-depth of a single lattice node, encompassing both Matter and Mirror sectors. This defines the Causality Limit derived in \cref{sec:fundamental_resonance}.
\begin{equation}
    N = 2 \times \nu = 16 + 16 = \mathbf{32}
\end{equation}

\subsubsection{The Bandwidth Limit (\texorpdfstring{$c_{eff} = 1$}{ceff})}
Because the lattice is discrete, information propagates via neighbor-to-neighbor updates. The maximum rate of causality is bounded by the lattice node spacing ($\ell$) divided by the update frequency ($\Delta$).

Consequently, \textbf{Lorentz Invariance} is identified not as a geometric axiom, but as the \textbf{Shannon Channel Capacity} ($C_{max}$) of the substrate.
The speed of light $c \equiv 1$ represents the maximum rate of causality: information can propagate exactly \textit{one lattice node spacing} ($\ell$) per \textit{one update cycle} ($\Delta$).

This enforces the constraints derived in \textbf{\cref{sec:fundamental_resonance_filter3}}: any signal attempting to propagate faster than the protocol allows results in temporal aliasing (Causal Violation) and is filtered by the vacuum's Causality Potential.